Supplemental Materials for \\ Helixer: Cross-species Gene Annotation of Large Eucaryotic Genomes Using Deep Learning

### 2020-05-25 paper/supplements.tex
some refs, some conventions, moving towards a hacked semi-dynamic ref for supplemental things
more semi-dynamic linking and refs
refs n cites
dynamic supplement and stuff

\documentclass{article}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{natbib}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

\begin{document}

@@

\author[1]{Felix Stiehler}
\author[1]{Alisandra Denton}
\author[1]{Marvin Steinborn}
\author[ \hspace{-1ex}]{Stephan Scholz}
\author[1]{Andreas Weber}
\affil[1]{Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}

@@

\newpage
\section{Model Evaluation by Species}
The following tables and plots show the prediction performance of Helixer for each species we worked with. The tables show a breakdown by basepair level accuracy, F1 score for each individual class and our two overall metrics, the Subgenic F1 and Genic F1. An F1 score of zero can indicate that the reference did not include that class.

@@

\renewcommand\thetable{}
\resizebox{\textwidth}{!}{%

@@

\clearpage
\begin{table}[!h]
\renewcommand\thetable{S1}
\resizebox{\textwidth}{!}{%
\centering
\begin{tabular}{@{}llllllll@{}}
\hline
Genome & Acc Overall & IG F1 & UTR F1 & CDS F1 & Intron F1 & Subgenic F1 & Genic F1 \\ [0.5ex]
\hline
tupaia\_belangeri & 0.8092 & 0.8584 & 0 & 0.732 & 0.7162 & 0.7167 & 0.7147 \\
tursiops\_truncatus & 0.8792 & 0.8984 & 0 & 0.8473 & 0.8556 & 0.8552 & 0.8535 \\
urocitellus\_parryii & 0.9391 & 0.9593 & 0.4454 & 0.8788 & 0.8948 & 0.8941 & 0.888 \\
ursus\_americanus & 0.9168 & 0.9455 & 0.3849 & 0.8449 & 0.8512 & 0.8508 & 0.8433 \\
ursus\_maritimus & 0.9257 & 0.9528 & 0.3279 & 0.8313 & 0.8528 & 0.852 & 0.8453 \\
vicugna\_pacos & 0.8516 & 0.886 & 0 & 0.7738 & 0.7945 & 0.7939 & 0.7923 \\
vombatus\_ursinus & 0.9315 & 0.9572 & 0.4186 & 0.8413 & 0.8494 & 0.8491 & 0.8425 \\
vulpes\_vulpes & 0.9349 & 0.9595 & 0.3522 & 0.8444 & 0.8616 & 0.8609 & 0.8541 \\
xenopus\_tropicalis & 0.8699 & 0.9175 & 0.317 & 0.7277 & 0.7438 & 0.7427 & 0.7289 \\
xiphophorus\_couchianus & 0.8575 & 0.5599 & 0 & 0.8585 & 0.9119 & 0.9052 & 0.8932 \\
xiphophorus\_maculatus & 0.8854 & 0.9156 & 0.5964 & 0.8368 & 0.871 & 0.8671 & 0.8439 \\
zonotrichia\_albicollis & 0.9216 & 0.9482 & 0.382 & 0.8601 & 0.8667 & 0.8663 & 0.857 \\
\hline
\end{tabular}}
\caption{Detailed results for all animal genomes.}
\end{table}

\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part1}}
\end{figure}
\begin{figure}[!h]
\renewcommand\thefigure{S1}
\centerline{\includegraphics[width=1.2\textwidth]{images/animals_f1_scores_part2}}
\caption{Subgenic F1 results for all animal genomes shown by phylogenetic position.}
\end{figure}

\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part1}}
\end{figure}
\begin{figure}[!h]
\renewcommand\thefigure{S2}
\centerline{\includegraphics[width=1.2\textwidth]{images/animals_differences_part2}}
\caption{Difference in Subgenic F1 to AUGUSTUS for all animal genomes shown by phylogenetic position.}
\end{figure}

@@

\renewcommand\thetable{S2}
\resizebox{\textwidth}{!}{%

@@

@@

\clearpage
\begin{figure}[!h]
\renewcommand\thefigure{S3}
\centerline{\includegraphics[width=1.0\textwidth]{images/plants_f1_scores}}
\caption{Subgenic F1 results for all plant genomes shown by phylogenetic position.}
\end{figure}
\begin{figure}[!h]
\renewcommand\thefigure{S4}
\centerline{\includegraphics[width=1.0\textwidth]{images/plants_differences}}
\caption{Difference in Subgenic F1 to AUGUSTUS for all plant genomes shown by phylogenetic position.}
\end{figure}

\clearpage
\section{Hyperparameters}

\begin{table}[!h]
\renewcommand\thetable{S3}
\centering
\begin{tabular}{@{}ll@{}}
\hline
Parameter & Value \\ [0.5ex]
\hline
LSTM Layers & 4 \\
LSTM Units per layer & 256 \\
Learning rate & 1e-3 \\
Batch Size & 52 \\
Optimizer & Adam \\
Clipnorm & 1.0 \\
Loss & Categorical Crossentropy \\
Float Precision & 32 bit \\
\hline
\end{tabular}
\caption{The hyperparameters that were used for the training of both the animal and plant models. For more implementation details please consult the source code.}
\label{suptab:params}
\end{table}

\begin{table}[!h]
\renewcommand\thetable{S4}
\centering
\begin{tabular}{@{}lll@{}}
\hline
Class & Animal Value & Plant Value \\ [0.5ex]
\hline
Intergenic & 0.7 & 0.3 \\
UTR & 1.6 & 1.0 \\
CDS & 1.2 & 0.9 \\
Intron & 1.2 & 0.3 \\
\hline
\end{tabular}
\caption{The class weights used during training. We found this parameter to be highly impactful for establishing a more balanced performance across the different classes. These are the only hyperparameters that are different for animals and plants. The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table [?]).}
\end{table}

\newpage
\section{Longer Sequence Input}
\label{sec:longer}
\begin{table}[!h]
\renewcommand\thetable{S5}
\centering
\begin{tabular}{@{}lll@{}}
\hline
Class & Input Length in bp\\ [0.5ex]
\hline
Mammalia & 200.000 \\
Non-avian Reptilia & 200.000 \\
Aves & 100.000 \\
Amphibia & 100.000 \\
Actinopterygii & 50.000 \\
Chondrichthyes & 50.000 \\
Actinistia & 50.000 \\
Ascidiacea & 20.000 \\
Insecta & 20.000 \\
\hline
\end{tabular}
\caption{Input sequence length for each phylogenetic class present in our animal dataset. As a rule of thumb, we found the input length should be proportional to the average gene length of the class while staying in the interval [20.000, 200.000]. If the N75 of a specific species was less than twice as high as the supposed sequence length it was lowered until either this criteria as met or a length of 50.000 was reached. }
\end{table}

@@

\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals1}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals2}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals3}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals4}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals5}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals6}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals7}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals8}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals9}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals10}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals11}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals12}}
\end{figure}
\begin{figure}[!h]
\renewcommand\thefigure{S5}
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals13}}
\caption{Effect of overlapping on the sequence positions wise bias ordered by descending N75 for all animals genomes individually.}
\end{figure}

@@

\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants1}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants2}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants3}}
\end{figure}
\begin{figure}[!h]
\renewcommand\thefigure{S6}
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_plants4}}
\caption{Effect of overlapping on the sequence positions wise bias ordered by descending N75 for all plant genomes individually.}
\end{figure}

\clearpage
\section{Training Data statistics}
\label{sec:training_data}
\begin{table}[!h]
\renewcommand\thetable{S6}
\centering
\begin{tabular}{@{}lll@{}}
\hline
& Animals & Plants\\ [0.5ex]
\hline
Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
\hline
\end{tabular}
\caption{The description of Table [?] applies here as well.}
\end{table}

\newpage
\section{Dilated CNN Baseline Network Search}
\begin{table}[!h]
\renewcommand\thetable{S7}
\centering
\begin{tabular}{@{}ll@{}}
\hline
Parameter & Search Space\\ [0.5ex]
\hline
Kernel Size & \{4, 8, 12, 16\}\\
Initial Filter Depth & \{32, 64, 96, 128\}\\
Number of Layers Before Doubling Filter Count & \{1, 2\}\\
Dilation Multiplier & \{2, 3\}\\
Number of Convolutional Layers & \{2, 3, 4, 5, 6, 7, 8\}\\
Number of Hidden Fully Connected (FC) Layers & \{0, 1, 2\}\\	
Dropout Used on FC and Final Conv Layer Output & \{0.0, 0.01, 0.1, 0.2, 0.3\}\\
Learning Rate & \{1e-3, 1e-4\}\\
\hline
\end{tabular}
\caption{Above are all parameters used during neural architecture search for the dilated CNN baseline. The same overall space was used for plant and animal data. We did, however, run multiple distinct searches that sometimes only operated over a subset of the given parameters. This was done as those seemed to be the most promising. We for example restricted the search space of the number of LSTM layers to the highest 3 values in later runs. Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with. Runs with a Genic F1 below 0.5 after 10 epochs were stopped and the overall maximum epoch was 15. The performances seemed to be leveling off well before that. The batch size used for almost all runs was 32, the dilation was capped to 81 and the size of each fully connected layer was fixed at 128. New parameters where chosen at random. The implementation and parameter space is very roughly based on \citep{gupta2017dilated} and can be found in the Helixer source code repository. In total, we trained 30 models for the animals and 48 with the plant data.}
\end{table}

\begin{table}[!h]
\renewcommand\thetable{S8}
\centering
\begin{tabular}{@{}lll@{}}
\hline
Parameter & Animals & Plants\\ [0.5ex]
\hline
Kernel Size & 16 & 16\\
Initial Filter Depth & 96 & 32\\
Number of Layers Before Doubling Filter Count & 2 & 2\\
Dilation Multiplier & 3 & 3\\
Number of Convolutional Layers & 6 & 8\\
Number of Hidden Fully Connected (FC) Layers & 2 & 0\\
Dropout Used on FC Output & 0.0 & 0.0\\
Learning Rate & 1e-4 & 1e-3\\
\hline
\end{tabular}
\caption{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes. The best animal model has circa 4.6 million parameters; the best plant model around 2.1 million.}
\end{table}

\section{Evaluation against RNAseq}

>>>>

\documentclass{article}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{natbib}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
\newgeometry{vmargin={20mm}, hmargin={25mm}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\def \tablescale{0.88}
\begin{document}

@@

\author[1]{Felix Stiehler}
\author[1]{Alisandra Denton}
\author[1]{Marvin Steinborn}
\author[ \hspace{-1ex}]{Daniela Dey}
\author[ \hspace{-1ex}]{Stephan Scholz}
\author[1]{Andreas Weber}
\affil[1]{Institue for Plant Biochemistry, Heinrich-Heine-University, Dusseldorf, D-40225, Germany}

@@

\newpage
\section{Data collection and preparation}
\label{ssec:data_prep}
% species and versions table
% meta data table?
% selecting species for RNAseq
\begin{figure}[!h]
\label{supfig:rna_select_animals}
\centerline{\includegraphics[width=1.0\textwidth]{images/RNAseq_animals_fit}}
\caption{Visualizes how the performance of Helixer (in consideration of Augustus' performance)
was used to select three animal species (larger points) for detailed RNAseq analysis. The 
species were chosen to have good, medium and poor performance; invertebrates (diamonds) were
not considered as the generalizability and recommended use of the Helixer animal model
currently only extends to vertebrates; candidate species were checked for the availibility of
stranded, paired-end RNAseq data and skipped if none was available.}
\end{figure}
\begin{figure}[!hbt]
\label{supfig:rna_select_plants}
\centerline{\includegraphics[width=0.5\textwidth]{images/RNAseq_plants}}
\caption{Visualizes how the performance of Helixer (in consideration of Augustus' performance)
was used to select three plant species (larger points) for detailed RNAseq analysis. The 
species were chosen to have good, medium and poor performance; algae (diamonds) were
not considered as the generalizability and recommended use of the Helixer animal model
currently only extends to higher plants; candidate species were checked for the availibility of
stranded, paired-end RNAseq data and skipped if none was available.}
\end{figure}
% RNAseq tested samples table
\resizebox{0.8\textwidth}{!}{%
\begin{tabular}{@{}lll|lll|lll|lll@{}}
Samples&HQ&S&Samples&HQ&S&Samples&HQ&S&Samples&HQ&S\\
\\
\textit{P. anubis}\\
SRS3122053&&&SRS3122094&\checkmark&\checkmark&SRS3270781&&&SRS3270847&&\\
SRS3122057&\checkmark&\checkmark&SRS3122111&&&SRS3270785&&&SRS3270853&&\\
SRS3122063&\checkmark&\checkmark&SRS3270734&&&SRS3270808&&&SRS3270858&&\\
SRS3122070&&&SRS3270735&&&SRS3270811&&&SRS3270862&&\\
SRS3122074&&&SRS3270737&&&SRS3270812&&&SRS3270863&&\\
SRS3122079&\checkmark&\checkmark&SRS3270738&&&SRS3270821&&&SRS3270870&&\\
SRS3122081&&&SRS3270746&&&SRS3270824&&&SRS3270872&&\\
SRS3122088&&&SRS3270751&&&SRS3270839&&&SRS3270880&&\\
SRS3122092&\checkmark&\checkmark&SRS3270763&&&SRS3270840&&&SRS819311&\checkmark&\checkmark\\
SRS3122093&&&SRS3270773&&&SRS3270841&&&SRS819317&\checkmark&\checkmark\\
\\
\textit{E. caballus}\\
ERS1517592&&&ERS2382444&&&ERS2681771&&&SRS2552997&\checkmark&\checkmark\\
ERS1517600&&&ERS2681720&\checkmark&\checkmark&ERS2681772&\checkmark&\checkmark&SRS2553004&\checkmark&\checkmark\\
ERS1517608&&&ERS2681730&&&ERS2681775&&&SRS2770963&&\\
ERS1517617&&&ERS2681733&&&ERS3563898&&&SRS2770972&&\\
ERS1517625&&&ERS2681734&&&SRS1102326&&&SRS3347177&&\\
ERS1517633&&&ERS2681740&&&SRS1102328&&&SRS3347182&&\\
ERS1517641&&&ERS2681742&&&SRS1102336&&&SRS4826859&&\\
ERS2382439&&&ERS2681750&&&SRS2158774&&&SRS4826867&&\\
ERS2382440&&&ERS2681754&&&SRS2158777&&&SRS4890460&\checkmark&\checkmark\\
ERS2382442&&&ERS2681758&&&SRS2158788&&&SRS5330332&\checkmark&\checkmark\\
ERS2382443&&&ERS2681770&&&SRS2552988&\checkmark&\checkmark&SRS5330340&&\\
\\
\textit{P. marinus}\\
SRS5315360&\checkmark&\checkmark&SRS5315363&\checkmark&&SRS5315366&\checkmark&&SRS5315369&\checkmark&\\
SRS5315361&\checkmark&\checkmark&SRS5315364&\checkmark&\checkmark&SRS5315367&\checkmark&\checkmark&&&\\
SRS5315362&\checkmark&\checkmark&SRS5315365&\checkmark&\checkmark&SRS5315368&\checkmark&\checkmark&&&\\
\\
\textit{M. esculenta}\\
SRS1608507&&&SRS1610089&\checkmark&\checkmark&SRS1610201&&&SRS946172&&\\
SRS1608510&\checkmark&\checkmark&SRS1610128&\checkmark&\checkmark&SRS1610205&\checkmark&\checkmark&SRS946227&&\\
SRS1610036&&&SRS1610150&&&SRS1610207&&&SRS946633&&\\
SRS1610041&&&SRS1610177&&&SRS1610208&&&SRS946650&&\\
SRS1610076&\checkmark&\checkmark&SRS1610181&&&SRS945853&&&SRS946703&&\\
SRS1610079&\checkmark&\checkmark&SRS1610187&&&SRS946078&&&SRS947897&&\\
SRS1610084&\checkmark&\checkmark&SRS1610197&&&SRS946138&&&&&\\
\\
\textit{M. truncatula}\\
ERS1556003&&&SRS1948297&&&SRS3700721&\checkmark&&SRS4425365&\checkmark&\\
ERS1556006&&&SRS1948300&&&SRS3700725&\checkmark&\checkmark&SRS4538246&&\\
ERS1556009&&&SRS1948304&\checkmark&&SRS4425341&&&SRS4538249&&\\
ERS1556012&&&SRS1948306&&&SRS4425343&\checkmark&&SRS5045319&&\\
ERS1556015&&&SRS1948309&&&SRS4425346&\checkmark&\checkmark&SRS5045322&&\\
ERS1556018&\checkmark&&SRS1948313&\checkmark&\checkmark&SRS4425347&\checkmark&\checkmark&SRS5045325&&\\
ERS1556021&&&SRS1948315&&&SRS4425350&\checkmark&&SRS5045328&&\\
ERS1556024&&&SRS3700712&\checkmark&&SRS4425353&&&SRS5045330&&\\
SRS1736395&\checkmark&\checkmark&SRS3700715&\checkmark&&SRS4425356&\checkmark&\checkmark&SRS5045334&&\\
SRS1736398&\checkmark&&SRS3700717&&&SRS4425357&&&SRS5045337&&\\
SRS1948293&&&SRS3700718&\checkmark&\checkmark&SRS4425364&\checkmark&&&&\\
\\
\textit{T. cacao}\\
SRS5244538&\checkmark&&SRS5244546&\checkmark&\checkmark&SRS5244554&&&SRS5244562&\checkmark&\checkmark\\
SRS5244539&&&SRS5244547&&&SRS5244555&&&SRS5244563&&\\
SRS5244540&&&SRS5244548&&&SRS5244556&\checkmark&\checkmark&SRS5244564&&\\
SRS5244541&&&SRS5244549&&&SRS5244557&\checkmark&&SRS5244565&\checkmark&\checkmark\\
SRS5244542&&&SRS5244550&&&SRS5244558&&&SRS5244566&&\\
SRS5244543&\checkmark&\checkmark&SRS5244551&&&SRS5244559&&&SRS5244567&&\\
SRS5244544&\checkmark&\checkmark&SRS5244552&&&SRS5244560&&&SRS5244568&&\\
SRS5244545&&&SRS5244553&\checkmark&\checkmark&SRS5244561&&&&&\\
\label{suptab:rnaseq_samples}
\caption{List of RNAseq samples that were selected for processing, whether they were of high quality (HQ), and 
finally, whether they were ultimately selected (S) for merging and use in coverage calculation.}
%

\section{Model Evaluation by Species}
\label{ssec:generalization}
The following tables and plots show the prediction performance of Helixer for each species we worked with. The tables show a breakdown by basepair level accuracy, F1 score for each individual class and our two overall metrics, the Subgenic F1 and Genic F1. An F1 score of zero can indicate that the reference did not include that class.

@@

%\renewcommand\thetable{}
\resizebox{\tablescale\textwidth}{!}{%

@@

\clearpage
\begin{table}[!h]
%\renewcommand\thetable{S1}
\resizebox{\tablescale\textwidth}{!}{%
\centering
\begin{tabular}{@{}llllllll@{}}
\hline
Genome & Acc Overall & IG F1 & UTR F1 & CDS F1 & Intron F1 & Subgenic F1 & Genic F1 \\ [0.5ex]
\hline
tupaia\_belangeri & 0.8092 & 0.8584 & 0 & 0.732 & 0.7162 & 0.7167 & 0.7147 \\
tursiops\_truncatus & 0.8792 & 0.8984 & 0 & 0.8473 & 0.8556 & 0.8552 & 0.8535 \\
urocitellus\_parryii & 0.9391 & 0.9593 & 0.4454 & 0.8788 & 0.8948 & 0.8941 & 0.888 \\
ursus\_americanus & 0.9168 & 0.9455 & 0.3849 & 0.8449 & 0.8512 & 0.8508 & 0.8433 \\
ursus\_maritimus & 0.9257 & 0.9528 & 0.3279 & 0.8313 & 0.8528 & 0.852 & 0.8453 \\
vicugna\_pacos & 0.8516 & 0.886 & 0 & 0.7738 & 0.7945 & 0.7939 & 0.7923 \\
vombatus\_ursinus & 0.9315 & 0.9572 & 0.4186 & 0.8413 & 0.8494 & 0.8491 & 0.8425 \\
vulpes\_vulpes & 0.9349 & 0.9595 & 0.3522 & 0.8444 & 0.8616 & 0.8609 & 0.8541 \\
xenopus\_tropicalis & 0.8699 & 0.9175 & 0.317 & 0.7277 & 0.7438 & 0.7427 & 0.7289 \\
xiphophorus\_couchianus & 0.8575 & 0.5599 & 0 & 0.8585 & 0.9119 & 0.9052 & 0.8932 \\
xiphophorus\_maculatus & 0.8854 & 0.9156 & 0.5964 & 0.8368 & 0.871 & 0.8671 & 0.8439 \\
zonotrichia\_albicollis & 0.9216 & 0.9482 & 0.382 & 0.8601 & 0.8667 & 0.8663 & 0.857 \\
\hline
\end{tabular}}
\label{suptab:generalization_animals}
\caption{Detailed results for all animal genomes.}
\end{table}

\begin{figure}[!h]
\label{supfig:phylogeny_f1_animals}
\centerline{\includegraphics[width=\textwidth]{images/animals_f1_scores_part1}}
\end{figure}
\begin{figure}[!h]
%\renewcommand\thefigure{S1}  % just commenting these out for now, incase I missed or miss-ordered anything in the text while making dynamic
\centerline{\includegraphics[width=\textwidth]{images/animals_f1_scores_part2}}
\caption{Subgenic F1 results for all animal genomes shown by phylogenetic position.}
\end{figure}

\begin{figure}[!h]
\label{supfig:phylogeny_differences_animals}
\centerline{\includegraphics[width=\textwidth]{images/animals_differences_part1}}
\end{figure}
\begin{figure}[!h]
%\renewcommand\thefigure{S2}
\centerline{\includegraphics[width=\textwidth]{images/animals_differences_part2}}
\caption{Difference in Subgenic F1 to AUGUSTUS for all animal genomes shown by phylogenetic position.}
\end{figure}

@@

%\renewcommand\thetable{S2}
\resizebox{\tablescale\textwidth}{!}{%

@@

\label{suptab:generalization_plants}

@@

\clearpage
\begin{figure}[!h]
\label{supfig:phylogeny_f1_plants}
%\renewcommand\thefigure{S3}
\centerline{\includegraphics[width=\tablescale\textwidth]{images/plants_f1_scores}}
\caption{Subgenic F1 results for all plant genomes shown by phylogenetic position.}
\end{figure}
\begin{figure}[!h]
\label{supfig:phylogeny_differences_plants}
%\renewcommand\thefigure{S4}
\centerline{\includegraphics[width=\tablescale\textwidth]{images/plants_differences}}
\caption{Difference in Subgenic F1 to AUGUSTUS for all plant genomes shown by phylogenetic position.}
\end{figure}

\clearpage
\section{Hyperparameters}
\label{sec:hyperparameters}
\begin{table}[!h]
%\renewcommand\thetable{S3}
\centering
\begin{tabular}{@{}ll@{}}
\hline
Parameter & Value \\ [0.5ex]
\hline
LSTM Layers & 4 \\
LSTM Units per layer & 256 \\
Learning rate & 1e-3 \\
Batch Size & 52 \\
Optimizer & Adam \\
Clipnorm & 1.0 \\
Loss & Categorical Crossentropy \\
Float Precision & 32 bit \\
\hline
\end{tabular}
\caption{The hyperparameters that were used for the training of both the animal and plant models. For more implementation details please consult the source code.}
\label{suptab:params}
\end{table}

\begin{table}[!h]
%\renewcommand\thetable{S4}
\centering
\begin{tabular}{@{}lll@{}}
\hline
Class & Animal Value & Plant Value \\ [0.5ex]
\hline
Intergenic & 0.7 & 0.3 \\
UTR & 1.6 & 1.0 \\
CDS & 1.2 & 0.9 \\
Intron & 1.2 & 0.3 \\
\hline
\end{tabular}
\caption{The class weights used during training. We found this parameter to be highly impactful for establishing a more balanced performance across the different classes. These are the only hyperparameters that are different for animals and plants. The exact weights were determined manually and partially reflect the difference in the class distributions of both datasets (see Table [?]).}
\label{suptab:classweights}
\end{table}

\newpage
\section{Longer Sequence Input}
\label{sec:longer}
\begin{table}[!h]
\label{suptab:prediction_lengths}
%\renewcommand\thetable{S5}
\centering
\begin{tabular}{@{}lll@{}}
\hline
Class & Input Length in bp\\ [0.5ex]
\hline
Mammalia & 200.000 \\
Non-avian Reptilia & 200.000 \\
Aves & 100.000 \\
Amphibia & 100.000 \\
Actinopterygii & 50.000 \\
Chondrichthyes & 50.000 \\
Actinistia & 50.000 \\
Ascidiacea & 20.000 \\
Insecta & 20.000 \\
\hline
\end{tabular}
\caption{Input sequence length for each phylogenetic class present in our animal dataset. As a rule of thumb, we found the input length should be proportional to the average gene length of the class while staying in the interval [20.000, 200.000]. If the N75 of a specific species was less than twice as high as the supposed sequence length it was lowered until either this criteria as met or a length of 50.000 was reached. }
\end{table}

@@

\subsection{Animals}
\newpage
\def \overlapscale{1.07}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals1}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals2}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals3}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals4}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals5}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals6}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals7}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals8}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals9}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals10}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals11}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_animals12}}
\end{figure}
\begin{figure}[!h]
%\renewcommand\thefigure{S5}
\centerline{\includegraphics[width=1.2\textwidth]{images/overlapping/montage_animals13}}
\caption{Effect of overlapping on the sequence positions wise bias ordered by descending N75 for all animals genomes individually.}
\label{supfig:overlapping_animals}
\end{figure}

@@

\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_plants1}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_plants2}}
\end{figure}
\begin{figure}[!h]
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_plants3}}
\end{figure}
\begin{figure}[!h]
%\renewcommand\thefigure{S6}
\centerline{\includegraphics[width=\overlapscale\textwidth]{images/overlapping/montage_plants4}}
\caption{Effect of overlapping on the sequence positions wise bias ordered by descending N75 for all plant genomes individually.}
\label{supfig:overlapping_plants}
\end{figure}

\clearpage
\section{Training Data statistics}
\label{sec:training_data}
\begin{table}[!h]
\label{suptab:genome_stats}
%\renewcommand\thetable{S6}
\centering
\begin{tabular}{@{}lll@{}}
\hline
& Animals & Plants\\ [0.5ex]
\hline
Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
\hline
\end{tabular}
\caption{The description of Table [?] applies here as well.}  % 2
\end{table}

\newpage
\section{Dilated CNN Baseline Network Search}
\label{sec:dcnn}
\begin{table}[!h]
\label{suptab:dcnn_xval}
%\renewcommand\thetable{S7}
\centering
\begin{tabular}{@{}ll@{}}
\hline
Parameter & Search Space\\ [0.5ex]
\hline
Kernel Size & \{4, 8, 12, 16\}\\
Initial Filter Depth & \{32, 64, 96, 128\}\\
Number of Layers Before Doubling Filter Count & \{1, 2\}\\
Dilation Multiplier & \{2, 3\}\\
Number of Convolutional Layers & \{2, 3, 4, 5, 6, 7, 8\}\\
Number of Hidden Fully Connected (FC) Layers & \{0, 1, 2\}\\	
Dropout Used on FC and Final Conv Layer Output & \{0.0, 0.01, 0.1, 0.2, 0.3\}\\
Learning Rate & \{1e-3, 1e-4\}\\
\hline
\end{tabular}
\caption{Above are all parameters used during neural architecture search for the dilated CNN baseline. The same overall space was used for plant and animal data. We did, however, run multiple distinct searches that sometimes only operated over a subset of the given parameters. This was done as those seemed to be the most promising. We for example restricted the search space of the number of LSTM layers to the highest 3 values in later runs. Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with. Runs with a Genic F1 below 0.5 after 10 epochs were stopped and the overall maximum epoch was 15. The performances seemed to be leveling off well before that. The batch size used for almost all runs was 32, the dilation was capped to 81 and the size of each fully connected layer was fixed at 128. New parameters where chosen at random. The implementation and parameter space is very roughly based on \citep{gupta2017dilated} and can be found in the Helixer source code repository. In total, we trained 30 models for the animals and 48 with the plant data.}
\end{table}

\begin{table}[!h]
\label{suptab:dcnn_params}
%\renewcommand\thetable{S8}
\centering
\begin{tabular}{@{}lll@{}}
\hline
Parameter & Animals & Plants\\ [0.5ex]
\hline
Kernel Size & 16 & 16\\
Initial Filter Depth & 96 & 32\\
Number of Layers Before Doubling Filter Count & 2 & 2\\
Dilation Multiplier & 3 & 3\\
Number of Convolutional Layers & 6 & 8\\
Number of Hidden Fully Connected (FC) Layers & 2 & 0\\
Dropout Used on FC Output & 0.0 & 0.0\\
Learning Rate & 1e-4 & 1e-3\\
\hline
\end{tabular}
\caption{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes. The best animal model has circa 4.6 million parameters; the best plant model around 2.1 million.}
\end{table}

\newpage
\section{Evaluation against RNAseq}
\label{sec:rnaseq_eval}
